---
title: "Lab 00 - Hello IDS!"
author: "Lucia ines"
date: "2025-09-30"
output: html_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

### Exercises 16-17

We set up the data frame. Edit the data below to include information about your team members.

```{r setup-tibble}
team_data <- tibble(
  name = c("Bonnie", "Caroline", "Klaus", "Lucia", "Lisa", "jeremy"),
  github_username = c("bonnie123", "caroC", "klausk", "luciafox", "lisalisa", "jeremy"),
  colour = c("blue", "red", "yellow", "pink", "violet", "green"),
  hobby = c("painting", "reading", "football", "reading", "knitting", "football"),
  favourite_number = c(7, 13, 3, 5, 11, 8)
)

team_data
```

### Exercise 18

We now work on visualising our data.

```{r bar-plot}
p1 <- ggplot(data = team_data, 
             mapping = aes(x = github_username, 
                           y = favourite_number,
                           fill = colour), 
             ) +
   geom_col() +
   scale_fill_identity()
p1
```

The plot shows a bar chart where each bar represents a team member (using their GitHub username). The height of the bar shows their favourite number.  The colour of each bar comes from the “colour” variable in the dataset.

### Exercises 19

Change the text of the plot title and axis labels.

```{r labelled-bar-plot}
p1 + labs(x = "Username ", 
          y = "Favourite Number",
          title = "Team Data")
```

The labs() function in ggplot2 is used to change the text labels in a plot. This includes the title of the plot, the x-axis label, the y-axis label, and sometimes legends. In this case, l chose tThe x-axis to say "username", y-axis to say "favourite number" and the title of the plot to "team data".

### Exercise 20

```{r new-bar-plot}
p2 <- ggplot(team_data, aes(x = hobby)) +
  geom_bar()
p2
```

Because if every person in the team has a different hobby, the bar chart will just show one bar per person. That doesn’t help us compare or find patterns — it's just a list. There’s no grouping or shared data to summarise. If lots of people had the same hobby, then the bar chart would show us which hobbies are more or less popular. A table might be better in this case. A table lets you list each team member and their hobby clearly, without pretending there’s a useful pattern. Another option might be a pie chart if hobbies were grouped or summarised, but only if some hobbies were shared by multiple people.
